Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-06 paper/main.tex
Added comment in Discussion
Several minor changes, and QA from start to 3.2

\label{sec:lengthextension}
\paragraph{Security.} The most important criteria for the SHA-3
candidates\cite{s_nistround2} is security. The full description of security
criteria are mentioned in \cite{s_fedreg}, however the most noteworthy criterion
in regards to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
family are vulnerable to this kind of attack, and thus requires
Tahoe to run the algorithm twice, as described by \citet{schneier}.

@@

\paragraph{Algorithm and implementation characteristics.} By this 3rd criteria,
\ac{NIST} emphasizes that algorithms with greater flexibility will be given
preference over other algorithms \cite{s_nistround2}. By flexible, they imply
the possibility of the function to run efficiently on a variety of platforms,
and to use parallelism and instruction set extensions. Another key point of a
flexible hash function is that it should have a simple and elegant design to
encourage understanding, analysis and design confidence.

\subsubsection{\ac{NIST} \ac{SHA}-3 \ac{API}}
\ac{NIST} required that every submission conform to a specified
\ac{API}\cite{s_fedreg}. The \ac{API} is specified at \cite{s_nistapi} and in
shortness it requires every candidate to implement four functions.

@@

The {\verb Init() } function basically sets up the internal state of the function, to
make it ready to start processing real data. The {\verb Update() } function is used to
provide the actual data to the hash function. The {\verb Final() } function should be
called when all necessary data has been given through {\verb Update() }, it will then
finish the hash function and provide the digest. The Hash() function basically a wrapper
for doing a call to the three other functions.

{\verb HashReturn } is an \emph{enum} representing that an operation either succeeded (0) or
went wrong. The {\verb BitSequence } is a representation of an array of byte fields and
{\verb DataLength } represents the size in bits of the data provided.

\subsubsection{The second round candidates}
The names of the second round candidates for the \ac{SHA}-3 \ac{NIST}
competition along with their principal submitters are listed in Table
\ref{tbl:sha3:candidates}\cite{s_nistround2}.

@@

\subsection{\ac{SUPERCOP}}
% TODO: Mention that some SUPERCOP candidates are missing UPDATE-functionality.
\ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
of cryptographic software \cite{s_supercop}. In relation to hash functions
\ac{SUPERCOP} measures the following:

@@

The Round 2 \ac{SHA}-3 candidates are all included in the toolkit, with a number
of different optimizations for each candidate. Optimizations range from
32-/64-bit specific implementations, the use of extended instruction sets such
as \ac{SSE2} and \ac{SSSE3}, optimizations for different number of cores and
others. The toolkit will also try different compiler optimizations to get the
best results possible for each function.
% TODO: Insert image to visualize the forest of different implementations.

The benchmarking results for the \ac{SUPERCOP} toolkit of the \ac{SHA}-3
candidates and \ac{SHA}-2 functions on a number of different plattforms and
architectures are available at
\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}
% TODO: Link to appendix with tahoe04 SUPERCOP results.

\subsubsection{\ac{SUPERCOP} \ac{API}}
\ac{SUPERCOP} specifies an \ac{API} which all submitted implementations must
conform to\cite{s_supercopapi}. This \ac{API} specifies the naming and
organization of files, and that a submission must include a function
{\verb crypto\_hash() } which is similar to the function {\verb Hash() } in the
\ac{NIST} \ac{SHA}-3 \ac{API}. The only difference being the naming of
data types and that the {\verb crypto\_hash() } function does not need an input
of how long the output of the hash function should be. Instead double or
different submissions are used for different output lengths.

@@

As seen in figure \ref{fig:tahoehashing}, the usage of secure hashing is
extensive in Tahoe, and has to be considered as a key part of the functionality,
and thus affecting the performance.

@@

A hash tree, also known as a \emph{Merkle tree}, is a type of data structure
which can be described as a tree with nodes that can verify all information
below in the hierarchy, as depicted in figure \ref{fig:tahoemerkletree}. This
enables Tahoe to verify small segments of a file at the time, and this can be
used for instance to start playing a movie file while it is still being
downloaded.

@@

A \textbf{capability} (or an URI) contains the encryption key, and a hash of the
\emph{\ac{UEB}}. The \ac{UEB} for each file is a data structure containing the
hash of the \ac{CEB}, the size of the file, any encoding parameters necessary to
perform the erasure decoding, and a hash of the plaintext and the encrypted
text. This is illustrated in figure \ref{fig:tahoeueb}.

@@

Direct calling of external C functions and methods is also supported, thus
making Cython an attractive way of wrapping external libraries written in C to
enable access from Python code.

@@

While the official \ac{NIST} \ac{SHA}-3 second round submissions contain both
a reference version and optimized versions of the candidates, it does not
appear to be updated since the submission deadline on September 15, 2009. They
also do not contain any benchmarking utility which could be used to decide on
the best implementation for a particular platform or architecture. \ac{SUPERCOP}
on the other hand, provides both, in combination with results stating which
compiler arguments gives the best performance. On the downside, the
\ac{SUPERCOP} \ac{API} only specifies the \texttt{crypto\_hash()} function and
not an \texttt{update()} function which Tahoe-\ac{LAFS} is dependent on using.

@@

We chose the fastest implementation of each candidate that conforms to the
following criteria:
\begin{enumerate}
  \item The implementation must have an \ac{API} defined in C.
  \item The implementation must include a working \verb update() function.
  \item The implementation must not use assembly that can not be made
    position independent by \ac{GCC}.
  \item The implementation should conform to the \acp{KAT} specified in the
    \ac{NIST} submission package, or later updates by the author.
\end{enumerate}

% TODO fix kilde
Based on these criteria, we choose the fastest possible implementation based
on our benchmark in \ac{SUPERCOP}. The relative speed of each implementation, in
relation to other implementation for the same and other candidates are
available at source.

@@

\input{tables/sha3_implementations}

The chosen implementations are shown in Table \ref{tbl:sha3:implementations}
and their compiler flags used, which are extracted from the \ac{SUPERCOP}
results can be seen in Table\ref{tbl:sha3:compilerflags}. It should be noted
that 32 in blake32 does not refer to the output length of the function, but the
fact that the algorithm is optimized for 32-bit usage. 512 in Keccackc512 also
does not mean 512-bit output, but 512 cycles in the internal function. The
relative speeds of the implementations chosen compared to the best version in
\ac{SUPERCOP} and relative speed of the benchmark compared to other candidates
can bee seen in Table \ref{tbl:sha3:speedrelative}. It should also be noted
that the fastest sha256 implementation in \ac{SUPERCOP} is
cryptopp\footnote{\url{http://www.cryptopp.com/}}, which is the one used by
Tahoe-\ac{LAFS}. For the most part we have found working versions of all
candidates in \ac{SUPERCOP}, however there are some exceptions.

@@

\paragraph{JH} seems to be broken in every implementation with regard to
criteria 2. For both the optimized versions found in \ac{SUPERCOP} and the
reference implementations in the \ac{NIST} submission package, Hash("foofoo")
will not yield the same result as two consecutive calls to Update("foo").
However the implementation in sphlib does fulfill this requirement. We have
also contacted the author, Thomas Pornin, which confirms this problem with the
official JH candidate and that the author has been made aware of the problem
but has not released any updated version yet.
% TODO: Fix hvordan vi referer til e-post?

@@

\paragraph{Simd vect128 implementation} does not by default conform to
criteria 4, because of a bug in the implementation. This has confirmed by the
author who also supplied us with a patch to correct the bug.

\paragraph{Blake32 \ac{SSSE3}} implementation also fail to wholly fulfill
criteria 4, however it does validate all test vectors that have a length 
in bits which is a multiple of 8. Since our Python wrappers only allow, and
only need, to provide data in full bytes we decided to keep this
implementation.

\paragraph{Shavite3} seems to have received an update\footnote{see
\url{http://www.cs.technion.ac.il/~orrd/SHAvite-3/}}. We have observed that
implementations of Shavite3 in \ac{SUPERCOP}-20101014 or later seems to fail to
verify the \ac{KAT}s, while the version we have used does. We believe this is
only because newer \ac{KAT}s will have to be generated, but since the author
has not done so we chose to use an older version of the algorithm.

% TODO: Add some lines that give reasons for the criteria above.

% sphlib, SUPERCOP, NIST

@@

\begin{itemize}
    \item What could have been done differently?
    \item Compare our results with general (supercop) results.
    \item Optimized candidates - which gain could be expected from more
    optimized candidates
    \item What if we had used another coding scheme? - does Cython lower the
    candidates performance?
    \item Implications of only testing immutable files? (mutable files also
    include directories)
    \item Adding a capability to support new hash functions in tahoe. (Our
    current way breaks backwards compatability)
\end{itemize}

>>>>

\label{sec:lengthextension}
\paragraph{Security.} The most important criterion for the SHA-3
candidates\cite{s_nistround2} is security. \citet{s_fedreg} lays out the full
description of security criteria, however the most noteworthy criterion
in regards to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
family are vulnerable to this kind of attack, and thus requires
Tahoe to run the algorithm twice, as described by \citet{schneier}.

@@

\paragraph{Algorithm and implementation characteristics.} By this 3rd criterion,
\ac{NIST} emphasizes that algorithms with greater flexibility will be given
preference over other algorithms \cite{s_nistround2}. By flexible, they imply
the possibility of the function to run efficiently on a variety of platforms,
and to use parallelism and instruction set extensions. Another key point of a
flexible hash function is that it should have a simple and elegant design to
encourage understanding, analysis and design confidence.

\subsubsection{\ac{NIST} \ac{SHA}-3 \ac{API}} \ac{NIST} required that every
submission should conform to a specified \ac{API}\cite{s_fedreg}. The \ac{API}
specification \cite{s_nistapi} states in shortness that it requires every
candidate to implement four functions:

@@

The {\verb Init() } function basically sets up the internal state of the function, to
make it ready to start processing real data. The {\verb Update() } function is used to
provide the actual data to the hash function. The {\verb Final() } function should be
called when all necessary data has been given through {\verb Update() }, it will then
finish the hash function and provide the digest. The {\verb Hash() } function
can be described as a wrapper for doing a call to the three other functions.

{\verb HashReturn } is an \emph{enum} representing that an operation either succeeded or
went wrong. The {\verb BitSequence } is a representation of an array of byte fields and
{\verb DataLength } represents the size in bits of the data provided.

\subsubsection{The second round candidates}
The names of the second round candidates for the \ac{SHA}-3 \ac{NIST}
competition along with their principal submitters are listed in Table
\ref{tbl:sha3:candidates} \cite{s_nistround2}.

@@

\subsection{\ac{SUPERCOP}}
\ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
of cryptographic software \cite{s_supercop}. In relation to hash functions,
\ac{SUPERCOP} measures the following:

@@

The Round 2 \ac{SHA}-3 candidates are all included in the toolkit, with a number
of different optimizations for each candidate. Optimizations range from
32-/64-bit specific implementations, the use of extended instruction sets such
as \ac{SSE2} and \ac{SSSE3}, optimizations for different number of cores and
others. The toolkit will also try different compiler optimizations to get the
best results possible for each function.

The benchmarking results for the \ac{SUPERCOP} toolkit of the \ac{SHA}-3
candidates and \ac{SHA}-2 functions on a number of different plattforms and
architectures are available at the webpage of eBACS
\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}.
%FIXME: Reference appendix if tahoe04 results end up there.

\subsubsection{\ac{SUPERCOP} \ac{API}}
\ac{SUPERCOP} specifies an \ac{API} which all submitted implementations must
conform to \cite{s_supercopapi}. This \ac{API} specifies the naming and
organization of files, and that a submission must include a function
{\verb crypto_hash() } which is similar to the function {\verb Hash() } in the
\ac{NIST} \ac{SHA}-3 \ac{API}. The only difference being the naming of
data types and that the {\verb crypto_hash() } function does not need an input
of how long the output of the hash function should be. Instead double or
different submissions are used for different output lengths.

@@

As seen in Figure \ref{fig:tahoehashing}, the usage of secure hashing is
extensive in Tahoe, and has to be considered as a key part of the functionality,
and thus affecting the performance.

@@

A hash tree, also known as a \emph{Merkle tree}, is a type of data structure
which can be described as a tree with nodes that can verify all information
below in the hierarchy, as depicted in Figure \ref{fig:tahoemerkletree}. This
enables Tahoe to verify small segments of a file at the time, and this can be
used for instance to start playing a movie file while it is still being
downloaded.

@@

A \textbf{capability} (or an URI) contains the encryption key, and a hash of the
\emph{\ac{UEB}}. The \ac{UEB} for each file is a data structure containing the
hash of the \ac{CEB}, the size of the file, any encoding parameters necessary to
perform the erasure decoding, and a hash of the plaintext and the encrypted
text. This is illustrated in Figure \ref{fig:tahoeueb}.

@@

Direct calling of external C functions and methods are also supported, thus
making Cython an attractive way of wrapping external libraries written in C to
enable access from Python code.

@@

While the official \ac{NIST} \ac{SHA}-3 second round submissions contain both
a reference version and optimized versions of the candidates, it does not
appear to be updated since the submission deadline on September 15, 2009. They
also do not contain any benchmarking utility which could be used to decide on
the best implementation for a particular platform or architecture. \ac{SUPERCOP}
on the other hand, provides both, in combination with results stating which
compiler arguments that gives the best performance. On the downside, the
\ac{SUPERCOP} \ac{API} only specifies the {\verb crypto_hash() } function and
not an {\verb update() } function which Tahoe-\ac{LAFS} is dependent on using.

@@

We chose the fastest implementation of each candidate that conforms to the
following criteria:
\paragraph{Criterion 1} The implementation must have an \ac{API} defined in C.
\paragraph{Criterion 2} The implementation must include a working {\verb update() } function.
\paragraph{Criterion 3} The implementation must not use assembly that can not be made
    position independent by \ac{GCC}.%FIXME: State the reason for this somewhere?
\paragraph{Criterion 4} The implementation should conform to the \acp{KAT} specified in the
    \ac{NIST} submission package, or later updates by the author.

\vspace{12pt}
Based on these criteria, we choose the fastest possible implementation based
on our benchmark in \ac{SUPERCOP}. The relative speed of each implementation, in
relation to other implementation for the same and other candidates are
available at source.
% TODO: Link to appendix with tahoe04 SUPERCOP results.

@@

\input{tables/sha3_implementations}
% FIXME: Insert image to visualize the forest of different implementations.

The chosen implementations are shown in Table \ref{tbl:sha3:implementations}.
Their compiler flags used, which are extracted from the \ac{SUPERCOP}
results, can be seen in Table \ref{tbl:sha3:compilerflags}.

It should be noted that \emph{32} in blake32 does not refer to the output length
of the function, but the fact that the algorithm is optimized for 32-bit usage.
Similarly, \emph{512} in keccakc512 does not mean 512-bit output, but 512 cycles in the
internal function.

The relative speeds of the implementations chosen compared to
the best version in \ac{SUPERCOP}, and relative speed of the benchmark compared
to other candidates, can bee seen in Table \ref{tbl:sha3:speedrelative}.
Worth mentioning is that the fastest SHA-256 implementation in \ac{SUPERCOP} is
cryptopp\footnote{\url{http://www.cryptopp.com/}}, which is the one used by
Tahoe-\ac{LAFS}. For the most part we have found working versions of all
candidates in \ac{SUPERCOP}, however there are some exceptions.

@@

\paragraph{JH} seems to be broken in every implementation with regard to
Criterion 2. For both the optimized versions found in \ac{SUPERCOP} and the
reference implementations in the \ac{NIST} submission package, Hash("foofoo")
will not yield the same result as two consecutive calls to Update("foo").
However the implementation in sphlib does fulfill this requirement. We have
also contacted the author, Thomas Pornin, which confirms this problem with the
official JH candidate and that the author has been made aware of the problem
but has not released any updated version yet.
% TODO: How to reference email?
% TODO: We have to write someting about sphlib itself

@@

\paragraph{Simd vect128 implementation} does not by default conform to Criterion
4, because of a bug in the implementation. This has been confirmed by the author
who also supplied us with a patch to correct the bug.

\paragraph{Blake32 \ac{SSSE3}} implementation also fail to completely fulfill
Criterion 4, however it does validate all test vectors that have a length
in bits which is a multiple of 8. Since our Python wrappers only allow (and
only need) to provide data in full bytes, we decided to keep this
implementation.

\paragraph{Shavite3} seems to have received an update\footnote{See
\url{http://www.cs.technion.ac.il/~orrd/SHAvite-3/}}. We have observed that
implementations of Shavite3 in \ac{SUPERCOP}-20101014 or later seems to fail to
verify the \ac{KAT}s, while the version we have used does. We believe this is
only because newer \ac{KAT}s will have to be generated, but since the author
has not done so, we chose to use an older version of the algorithm.

@@

% TODO: Mention that some SUPERCOP candidates are missing UPDATE-functionality.
% What could have been done differently?
% Compare our results with general (supercop) results.
% Optimized candidates - which gain could be expected from more optimized
%   candidates?
% What if we had used another coding scheme? - does Cython lower the
%   candidates performance?
% Implications of only testing immutable files? (mutable files also
%   include directories)
% Adding a capability to support new hash functions in tahoe. (Our
%   current way breaks backwards compatability)
% Time gain of removing ``double'' SHA-256. Double only in digest()
